A boundary-integral framework to simulate \\ viscous erosion of a porous medium

### 2018-07-03 docs/methods/erosion_methods.tex
fixed typo Lukas found in equation (10)
Making edits for the erosion timescale (Ref 3, comment 1)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We consider flows that are confined by a solid wall $\Gamma$ that encloses $M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  Therefore, the boundary of the fluid domain is $\bd\Omega = \Gamma \cup \gamma$. On each body, $\nn$ and $\ss$ denote the unit normal and unit tangent vectors respectively.  Throughout the paper, we adopt the convention that $\ss$ points in the counterclockwise direction and $\nn$ points out of the fluid domain (i.e.~into the bodies). Neglecting inertial forces, the dynamics of the fluid is fully determined by the position of the bodies $\xx_\ell(s,t) \in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given $\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the fluid viscosity.  A scaling constant often appears in the erosion model in~\eqref{eqn:erosionModel}, but in the Stokes limit, the constant only sets the time scale, so we have taken it to be one.  For the sake of numerical stability, we will modify the erosion model in Section~\ref{sec:thetaL}.

@@

The deformation tensor of the double-layer potential~\eqref{eqn:velocityDLP} at an interior point $\xx \in \Omega$ is~\cite{qua-bir2014a}
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) \, ds_\yy \, .
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends to $\xx_0 \in \bd\Omega$.  The limiting value is
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0)\, , \quad \xx_0 \in \bd\Omega \, ,
\end{align} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right] \, .
  \label{eqn:shearStressJump}
\end{align}
Next, the deformation tensor of the Stokeslets and rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right) \, .
  \end{aligned}
\end{equation}
Combining these deformation tensors, the shear stress at $\xx_0 \in \bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss \, .
  \label{eqn:shearStressLP}
\end{align}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We consider flows that are confined by a solid wall $\Gamma$ that encloses $M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  Therefore, the boundary of the fluid domain is $\bd\Omega = \Gamma \cup \gamma$. On each body, $\nn$ and $\ss$ denote the unit normal and unit tangent vectors respectively.  Throughout the paper, we adopt the convention that $\ss$ points in the counterclockwise direction and $\nn$ points out of the fluid domain (i.e.~into the bodies). Neglecting inertial forces, the dynamics of the fluid is fully determined by the position of the bodies $\xx_\ell(s,t) \in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given $\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model} .
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface. The parameters in this system are $\mu$ the fluid viscosity and $\CE$ the mater-dependent erosion constant.
%A scaling constant often appears in the erosion model in~\eqref{eqn:erosionModel}, but in the Stokes limit, the constant only sets the time scale, so we have taken it to be one.  For the sake of numerical stability, we will modify the erosion model in Section~\ref{sec:thetaL}.

@@

The deformation tensor of the double-layer potential~\eqref{eqn:velocityDLP} at an interior point $\xx \in \Omega$ is~\cite{qua-bir2014a}
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \, ds_\yy \, .
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends to $\xx_0 \in \bd\Omega$.  The limiting value is
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0)\, , \quad \xx_0 \in \bd\Omega \, ,
\end{align} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right] \, .
  \label{eqn:shearStressJump}
\end{align}
Next, the deformation tensor of the Stokeslets and rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right) \, .
  \end{aligned}
\end{equation}
Combining these deformation tensors, the shear stress at $\xx_0 \in \bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss \, .
  \label{eqn:shearStressLP}
\end{align}
